Draft

### 2017-08-28 Thesis/Draft.tex
Update SLR section

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[lined,boxed]{algorithm2e}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}

@@

Another algorithm for the thresholding bandit problem is introduced in
Zhong et al. (2017). Indeed, they discuss several variations of their
algorithm which extend their central idea to applications in which
several samples need to be pulled in a parallel and asynchronous
fashion. An easy example known from the multi-armed bandit literature
(SOURCE????!!??!?!?!) is a website where different versions are served
to customers and each version is an arm in the bandit setting. If the
feedback that is observed by the algorithm is whether or not a user
makes a purchase, then it is likely that a new user arrives before the
feedback for the previous user has been observed. The algorithm has to
pull the next arm before the feedback of the previous arm has been
observed. With their variants the authors approach this problem of
having to take asynchronous and parallel actions. But their algorithm
can also be reduced to a simpler version that can be compared directly
with the problem described in Locatelli et al. (2016) where we assume
the problem of the feedback lag away.

This simpler version is the anytime parameter-free version of their
Empirical Variance Guided Algorithm (EVT\(_{pf}\)). The EVT\(_{pf}\)
algorithm is a smart extension of the APT algorithm to an approach that
takes the empirical variance of the arms into account and deduces an
upper bound through Bernstein bounds.

In order to describe their algorithm, we adapt a formulation used in
Zhong et al. (2017) to describe the index used by the algorithms in a
more general fashion. Write the index \(B_i(t)\) of arm \(i\) at round
\(t\) as previously as

where \(D_i(t)\) is some measure of the distance between the mean
\(\mu_i\) or the distribution \(\nu_i\) of the arm \(i\) from the
threshold \(\tau\), while \(S_i(t)\) is some confidence value. Thus, for
the APT algorithm, if we let \(\epsilon = 0\) as in Zhong et al. (2016),
we have

\[
D_i^{APT}(t) = \hat{\Delta}_i(t) = |\hat{\mu}_i(t) - \tau| \quad \land \quad S_i^{APT}(t) = \sqrt{T_i(t)}
\]

@@

\[
D_i^{EVT}(t) = \hat{\Delta}_i(t) = |\hat{\mu}_i(t) - \tau| \quad \land \quad S_i^{EVT}(t) = \Big(\frac{1}{T_i(t) + \sqrt{\frac{1}{T_i(t)}}\hat{\sigma}_{i}(t)}\Big)^{-1}
\]

or equivalently as

\[
D_i^{EVT}(t) = \sqrt{\hat{\sigma}^2_{i}(t) + \hat{\Delta}_i(t)} - \hat{\sigma}_i(t) \quad \land \quad S_i^{EVT} = \sqrt{T_i(t)}
\]

We see that in both formulations either the distance term \(D_i\) or the
confidence term \(S_i\) between the two algorithms is equivalent. While
the empirical variance is a natural confidence measure -- and thus makes
sense in the \(S_i^{EVT}\) part in the first equation -- it also
describes the shape of the distribution and thus is also a measure of
how far the distribution (in contrast to just the mean) is away from the
threshold, and thus can belong to \(D_i^{EVT}\). We will focus on the
latter formulation for the remaining paper.

In comparison with the APT algorithm, the EVT\(_{pf}\) algorithm adds
the empirical variance into the index to measure how well we can expect
the distribution (and the mean) to be seperated from the threshold; it
serves as a crude measure of distance between threshold and
distribution. We note that for a Bernoulli distributed random variable,
\(D_i^{EVT}\) is decreasing as the mean approaches the threshold.
Similar should hold in general for bounded variables as the variance and
the mean are linked by the distance of the mean to the bound of the
support (in contrast to, for example, a normal distributed random
variable where a change in the mean does not correspond to a change in
the variance). This is important as the bounds of the EVT\(_{pf}\)
algorithm only hold for random variables bounded in \([0,1]\).

@@

The authors approach the proof very similarly as Locatelli et al. (2016)
in the proof of their lower bound. They define favorable events that
hold with large probability, make sure the same separation condition
holds for them given that each arm is pulled often enough. Additionally
to a favorable event that bounds the deviation between sample mean and
population mean, they define a second event on which the sample standard
deviation is close to the population standard deviation. This is bounded
in large probability by a Bernstein inequality. The union of the two
events then serves as the favorable event in the remainder of their
proof.

Given a lower bound that is also adapted to the new complexity
\(H^{\text{EVT}_{pf}}\), the authors show that the EVT\(_pf\) algorithm
is optimal. Furthermore, they compare their algorithms in experiments
against the APT algorithm. The experiments use data generated from
uniform distributions. Each experiment compares 100 arms with variances
picked at random from a range of possible variances. Over three
experiments, the range of variances is varied from small to large, where
the authors expect their algorithm to show the largest advantage
compared to APT on the problem with small variances. Interestingly, they
report the largest advantage of their variance based algorithms in
comparison to APT when the variances are very small; given they only
consider problems bounded on a support of \([0,1]\), \(\sigma_i \ll 1\).
In this case the problem complexity \(H^{\text{EVT}_{pf}}\) can be much
smaller than \(H^{APT}\) -- the EVT algorithm profits more from
distributions that are simple to estimate in terms of their variance
than the APT algorithm does.

When comparing the two complexities for example for Bernoulli random
variables against some fixed threshold, we see that the smaller
complexity of the EVT\(_{pf}\) algorithm implies in particular that
EVT\(_{pf}\) could be able to cope with distributions of parameter
\(\mu\) close to \(\tau\) much better than APT.

@@

{\centering \includegraphics[width=0.8\linewidth]{Draft_files/figure-latex/Hevt_Hapt_Comparison-1}

@@

\caption{This graph compares the complexity terms used in the APT, AugUCB, and EVT algorithms respectively. We observe that as the mean of the arm's distribution is closer to the threshold, the complexity of the problem increases much less rapidly for the EVT algorithm. Thus, the EVT should be able to cope with arms close to the threshold better than the APT algorithm. The complexity displayed for the AugUCB is an upper bound on the complexity used in the upper bound of Mukherjee et al. (2017).}\label{fig:Hevt_Hapt_Comparison}
\end{figure}

\subsection{\texorpdfstring{Likelihood Ratio for Exponential Family
Distributions
\label{sec:LikelihoodRatioForExponentialFamily}}{Likelihood Ratio for Exponential Family Distributions }}\label{likelihood-ratio-for-exponential-family-distributions}

We now present an approach that builds on the idea of the likelihood
ratio statistic for hypothesis tests. We first present the general idea
of likelihood ratio based tests, and then introduce our new algorithm
for the thresholding bandit problem.

\subsubsection{Likelihood Ratio-based Hypothesis
Tests}\label{likelihood-ratio-based-hypothesis-tests}

In a standard parametric testing problem, we have drawn an iid sample
\(x = (x_1, ..., x_T)\), where each \(x_i\) is a realization from the
random variable \(X_i\). We assume that \(X_i\) follows some probability
distribution \(Q(\theta)\) paramaterized by a vector \(\theta\). Let
\(q(x;\theta)\) be the density function corresponding to the
distribution. In a standard hypothesis test, we test whether an
alternative parameterization is more likely than the parameterization
given by the null hypothesis. That is, we test

@@

In order to assess whether the \(H_0\) hypothesis can be rejected, one
computes the likelihood \(L(x|\theta)\) of the observed sample under
both parameterizations:
\(L_{H_0} = L(x|\theta = \theta_{H_0}) = \prod_{i=1}^T q(x_i|\theta_{H_0})\)
and
\(L_{H_1} = L(x|\theta = \theta_{H_1}) = \prod_{i=1}^T q(x_i|\theta_{H_1})\).
The likelihood ratio test statistic is then given as

\[
LR(\theta_{H_0}|| \theta_{H_1}) = \frac{L_{H_0}}{L_{H_1}} = \prod_{i=1}^T\frac{q(x_i|\theta_{H_0})}{ q(x_i|\theta_{H_1})}.
\] The intuition behind the test of course being that, if the true
parameterization is \(H_1\) and not \(H_0\), then the likelihood
\(L_{H_1}\) should be larger than \(L_{H_0}\). A small enough likelihood
ratio would thus lead us to a rejection of the null hypothesis, while a
large likelihood ratio is evidence in favor of the null hypothesis.

By considering \(lr = -2\log(LR(\theta_{H_0}|| \theta_{H_1}))\), one has
a statistic that is \(\chi^2\) distributed with degrees of freedom that
depend on the null and alternative hypothesis. Assessing whether the
null hypothesis can be rejected with large confidence thus becomes
straightforward.

\subsubsection{Likelihood Ratio-based Thresholding
Bandit}\label{likelihood-ratio-based-thresholding-bandit}

The simple concept of the likelihood-ratio test serves as foundation for
a new sampling strategy that can be applied to the thresholding bandit
problem. In contrast to the cumulative regret problem, or even the best
arm identification problem, we have some sort of null hypothesis for the
distributions in the thresholding bandit problem. In case of no
information and no observed samples, the null hypothesis might simply be
that all distributions have a mean that equals the threshold. By
observing samples, we then want to test whether the mean of the
distributions is different from the threshold \(\tau\) in either
direction. Once we are sure that there is a difference, we do not want
to continue to test. Instead, we would like to focus our sampling effort
on a different arm where we are not as sure. The likelihood ratio
quantifies how ``significant''" of a difference there is between the
threshold \(\tau\) and the current sample mean \(\hat{\mu}_i(t)\) of an
arm. Given that we plug in the sample mean, that is, the maximum
likelihood estimator for the true mean for many distributions, the
likelihood ratio will be smaller than 1 and decrease in sample size even
if the threshold and sample mean are constant.

Due to these characteristics, the likelihood ratio can serve as a
natural index for sampling strategies. Thus write as index

\[
B_i(t) = LR(\tau|| \hat{\mu}_i(t)) = \prod_{i=1}^T\frac{q(x_i|\tau)}{ q(x_i|\hat{\mu}_i(t))}
\]

and pull at round \(t\) the arm \(i^* = \arg \max_i B_i(t)\). As we
described in the previous paragraph, we have \(B_i(t) \leq 1\),
\(t \uparrow \rightarrow B_i(t) \downarrow\), and
\(|\hat{\mu}_i(t) - \tau| \uparrow \rightarrow B_i(t) \downarrow\).
Consequently, by choosing the arm that maximizes \(B_i(t)\), we choose
an arm for which we either have not collected many samples, or where the
sample mean is close to the threshold \(\tau\). In this sense, a
strategy based on the likelihood ratio is of course very reminiscient of
the index employed by the APT algorithm. Indeed, assuming the samples
are normally distributed with the same variance for the distribution
\(\nu_i\) under both \(H_0\) and \(H_1\), then
\(-\log(LR(\tau||\hat{\mu}_i(t)))\) equals the index in APT. Why then
would one use the likelihood based strategy?

Given that the APT index and the likelihood-ratio based index are only
equivalent for normally distributed random variables with fixed standard
deviation, the new strategy promises efficiency improvements whenever
the true distribution is known and different from the normal
distribution. The most important example are Bernoulli arms with small
means as discussed before. Here, it is also easy to assume that the true
distribution is known upfront. Similarly, the likelihood ratio statistic
can be used for other univariate distributions of the exponential
family. But we also propose a version for the Normal distribution with
unknown mean.

We now focus on the case where the distributions
\(\nu_{\theta_1}, ..., \nu_{\theta_K}\) are all the same canonical
exponential family of distributions with scalar parameter
\(\theta_i \in \Theta\). The probability density function of the
absolutely continuous distribution \(\nu_{\theta_i}\) with regard to a
dominating measure \(\rho\) on \(\mathbb{R}\) is given by
\(q(x|\theta) = h(x) \exp(T(x)\theta - b(\theta))\). As described in
Cappé et al. (2013), citing Lehmann and Casella (1998), the derivative
\(b'(\theta)\) of \(b(\theta)\) is an increasing continuous function
such that the expected value \(\mathbb{E}(\nu_{\theta}) = b'(\theta)\)
for all \(\theta \in \Theta\). Given that \(b'(x)\) is strictly convex,
we have a one-to-one relation between \(b'(\theta)\) and
\(\mathbb{E}(\nu_{\theta})\) through the continuous inverse
\(b'^{-1}(\theta)\). Consequently, one can just as well parameterize the
distribution its mean \(\mu = \mathbb{\nu_{\theta}}\) and write
\(\nu_{b'^{-1}(\mu)}\) with the probability density function

@@

As we will see now, this statistic is easily derived for many specific
distributions.

\paragraph{Example 1: Bernoulli
Distribution}\label{example-1-bernoulli-distribution}

@@

\[
-\log(LR(\mu_0||\mu_1)) = n(\log\Big(\frac{1-\mu_0}{1-\mu_1}\Big)) - \Big(\log(\frac{\mu_0}{1-\mu_0}) - \log(\frac{\mu_1}{1-\mu_1})\Big)\sum_{i=1}^{n}x_i = n \Big( (1-\hat{\mu})\log\Big(\frac{1-\mu_1}{1-\mu_0}\Big) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}) \Big)
\]

We observe that, if we let \(\mu_0 = \tau\) and \(\mu_1 = \hat{\mu}\),
then

@@

\paragraph{Example 2: Poisson
Distribution}\label{example-2-poisson-distribution}

@@

We observe that, if we let \(\mu_0 = \tau\) and \(\mu_1 = \hat{\mu}\),
then

@@

\paragraph{\texorpdfstring{Example 3: Normal distributions with known
variance
\(\sigma^2\)}{Example 3: Normal distributions with known variance \textbackslash{}sigma\^{}2}}\label{example-3-normal-distributions-with-known-variance-sigma2}

@@

And again, if we let \(\mu_1 = \hat{\mu}\) and \(\mu_0 = \tau\), then

@@

\paragraph{\texorpdfstring{Example 4: Gamma distributions with known
shape parameter
\(\alpha\)}{Example 4: Gamma distributions with known shape parameter \textbackslash{}alpha}}\label{example-4-gamma-distributions-with-known-shape-parameter-alpha}

@@

\[
-\log(LR(\mu_0||\mu_1)) = n\alpha(-\log(\frac{\alpha}{\mu_0}) - \log(\frac{\mu_1}{\alpha})) - (-\frac{\alpha}{\mu_0} + \frac{\alpha}{\mu_1})\sum_{i=1}^{n}x_i = n\alpha \Big(\log\big(\frac{\mu_0}{\mu_1}\big) - \frac{\hat{\mu}}{\mu_1} + \frac{\hat{\mu}}{\mu_0}\Big).
\] And again, if we let \(\mu_1 = \hat{\mu}\) and \(\mu_0 = \tau\), then

@@

Given these results for the exponential family of distributions, we can
now define the SLR (Simple Likelihood Ratio) algorithm.

\subsubsection{1D-SLR Algorithm for One-Dimensional Exponential Family
of
Distributions}\label{d-slr-algorithm-for-one-dimensional-exponential-family-of-distributions}

>>>>

\usepackage[retainorgcmds]{IEEEtrantools}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{hyperref}
\usepackage[lined,boxed]{algorithm2e}
\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newcommand{\KL}{\,\text{KL}}
\newcommand{\der}{\,\text{d}}
\newcommand*{\Alignyesnumber}{\refstepcounter{equation}\tag{\theequation}}%

@@

Another algorithm for the thresholding bandit problem employing
empirical variance estimates is introduced in Zhong et al. (2017).
Indeed, the authors discuss several variations of their algorithm which
extend their central idea to applications in which several samples need
to be pulled in a parallel and asynchronous fashion. An standard example
known from the multi-armed bandit literature (see for example Li et al.
(2010)) is a website for which different versions are served to visitors
and each version is an arm in the bandit setting. If the feedback that
is observed by the algorithm is whether or not a user makes a purchase,
then it is likely that a new user arrives before the feedback for the
previous user has been observed. The algorithm has to pull the next arm
before the feedback of the previous arm has been observed. With their
variants the authors approach this problem of having to take
asynchronous and parallel actions. But their algorithm can also be
reduced to a simpler version that can be compared directly with the
problem described in Locatelli et al. (2016) where the feedback lag is
assumed to be zero.

This simpler version is the parameter-free version of their Empirical
Variance Guided Algorithm (EVT\(_{pf}\)). The EVT\(_{pf}\) algorithm is
an extension of the APT algorithm that takes the empirical variance of
the arms into account and deduces an upper bound through empirical
Bernstein bounds in a proof following the model of the one described in
Locatelli et al. (2016).

In order to describe their algorithm, we adapt a formulation used in
Zhong et al. (2017) to describe the index used by the algorithms in a
more general fashion. Write the index \(B_i(t+1)\) of arm \(i\) at round
\(t\) as previously as

\[
B_i(t+1) = S_i(t) D_i(t),
\]

where \(D_i(t)\) is some measure of the distance between the mean
\(\mu_i\) or the distribution \(\nu_i\) of arm \(i\) from the threshold
\(\tau\), while \(S_i(t)\) is some confidence value measuring how
confident we can be in \$. Thus, for the APT algorithm, if we let
\(\epsilon = 0\) as in Zhong et al. (2016), we have

\[
D_i^{APT}(t) = \hat{\Delta}_i(t) = |\hat{\mu}_i(t) - \tau| \quad \text{and} \quad S_i^{APT}(t) = \sqrt{T_i(t)}
\]

@@

\[
D_i^{EVT_{pf}}(t) = \sqrt{\hat{\sigma}^2_{i}(t) + \hat{\Delta}_i(t)} - \hat{\sigma}_i(t) \quad \text{and} \quad S_i^{EVT_{pf}} = \sqrt{T_i(t)}
\]

We see that both algorithms employ the same confidence term. The
empirical variance describes the shape of the distribution and thus is a
measure of how far the distribution (in contrast to just the mean) is
away from the threshold. It consequently makes sense to have it appear
in \(D_i^{EVT}\). And in contrast to the APT algorithm, the EVT\(_{pf}\)
algorithm consequently takes more information about the arms into
account than just their mean. As mentioned previously, this should
especially be valuable whenever arms have indeed different variances.
Similar (very roughly) to how the Kullback-Leibler divergence measures
distances between distributions, the \(D_i^{EVT}\) including the
empirical variance serves as a crude measure of distance between
threshold and distribution. We note that for a Bernoulli distributed
random variable, \(D_i^{EVT}\) is always decreasing as the mean
approaches the threshold (compare the figure 2). Similar should hold in
general for bounded variables as the variance and the mean are linked by
the distance of the mean to the bound of the support (in contrast to,
for example, a normal distributed random variable where a change in the
mean does not correspond to a change in the variance). This is important
as the bounds of the EVT\(_{pf}\) algorithm only hold for random
variables bounded in \([0,1]\). In contrast to the AugUCB, this
restriction is not just on the variance of the arm but explicitly on the
reward itself in the current version that builds on empirical Bernstein
bounds as shown in Maurer and Pontil (2009).

@@

The authors approach the proof very similarly as Locatelli et al. (2016)
in the proof of their lower bound. They define favorable events that
hold with large probability, and ensure that the same separation
condition as in Locatelli et al. holds for them given that each arm is
pulled sufficiently often. Additionally to a favorable event that bounds
the deviation between sample mean and population mean, they define a
second event on which the sample standard deviation is close to the
population standard deviation. This is bounded in large probability by
an empirical Bernstein inequality (Maurer and Pontil, 2009). The union
of the two events then serves as the favorable event in the remainder of
their proof.

Given a lower bound adapted to the new complexity
\(H^{\text{EVT}_{pf}}\), the Zhong et al. (2017) show that the
EVT\(_pf\) algorithm is optimal in that its upper bound matches the
lower bound. Furthermore, they compare their algorithm in experiments
against the APT algorithm. The experiments use data generated from
uniform distributions. Each experiment compares 100 arms with variances
picked at random from a range of possible variances. Over three
experiments, the range of variances is varied from small to large, where
the authors expect their algorithm to show the largest advantage
compared to APT on the problem with small variances. Interestingly, they
report the largest advantage of their variance based algorithms in
comparison to APT when the variances are very small; given they only
consider problems bounded on a support of \([0,1]\), \(\sigma_i \ll 1\).
In this case the problem complexity \(H^{\text{EVT}_{pf}}\) can be much
smaller than \(H^{APT}\) -- the EVT algorithm profits more from
distributions that are simple to estimate in terms of their variance
than the APT algorithm does.

When comparing the two complexities for example for Bernoulli random
variables against some fixed threshold, we see that the smaller
complexity of the EVT\(_{pf}\) algorithm implies in particular that
EVT\(_{pf}\) could be able to cope with distributions of parameter
\(\mu\) close to \(\tau\) much better than APT (compare figure 3).

@@

{\centering \includegraphics[width=0.7\linewidth]{Draft_files/figure-latex/Hevt_Hapt_Comparison-1}

@@

\caption{Comparing the complexity terms used in the APT, AugUCB, and EVT algorithms, we observe that the problem complexity of the EVT algorithm increases much less rapidly than for the other algorithms. Thus, the EVT should be able to cope with arms close to the threshold better than the APT algorithm. The complexity displayed for the AugUCB is an upper bound on the complexity used in the upper bound of Mukherjee et al. (2017).}\label{fig:Hevt_Hapt_Comparison}
\end{figure}

\subsection{\texorpdfstring{Likelihood Ratio for Sequential Sampling
\label{sec:LikelihoodRatioForExponentialFamily}}{Likelihood Ratio for Sequential Sampling }}\label{likelihood-ratio-for-sequential-sampling}

We now present a new sequential sampling algorithm for the fixed budget
thresholding bandit problem that builds on the idea of the likelihood
ratio statistic for hypothesis tests. We first shortly describe the
general idea of likelihood ratio based tests, to then introduce this new
algorithm for the thresholding bandit problem, which we call Simple
Likelihood Ratio (SLR) algorithm.

\subsubsection{\texorpdfstring{Likelihood Ratio-based Hypothesis Tests
\label{sec:LRTests}}{Likelihood Ratio-based Hypothesis Tests }}\label{likelihood-ratio-based-hypothesis-tests}

In a standard parametric testing problem, we have drawn an iid sample
\(x = (x_1, ..., x_T)\), where each \(x_i\) is a realization from the
random variable \(X_i\). We assume that \(X_i\) follows some probability
distribution \(Q(\theta)\) paramaterized by a vector \(\theta\). Let
\(q(x;\theta)\) be the density function corresponding to the
distribution. We would like to test whether an alternative
parameterization is more likely than the parameterization given by the
null hypothesis. That is, we test

@@

In order to assess whether the \(H_0\) hypothesis can be rejected, one
computes the likelihood \(L(x|\theta)\) of the observed sample under
both parameterizations:
\(L_{H_0} = L(x|\theta = \theta_{H_0}) = \prod_{i=1}^T q(x_i|\theta_{H_0})\)
and
\(L_{H_1} = L(x|\theta = \theta_{H_1}) = \prod_{s=1}^T q(x_s|\theta_{H_1})\).
The likelihood ratio test statistic is then given as

\[
LR_T(\theta_{H_0}|| \theta_{H_1}) = \frac{L_{H_0}}{L_{H_1}} = \prod_{s=1}^T\frac{q(x_s|\theta_{H_0})}{ q(x_s|\theta_{H_1})}.
\]

and is used in the optimal test of two hypotheses as shown by the
Neyman-Pearson lemma (see chapter 11, Theorem 11.7.1 in Cover and
Thomas, 2006).

The intuition behind the test of course being that, if the true
parameterization is \(H_1\) and not \(H_0\), then the likelihood
\(L_{T,H_1}\) should be larger than \(L_{T,H_0}\). A small enough
likelihood ratio would thus lead us to a rejection of the null
hypothesis, while a large likelihood ratio is evidence in favor of the
null hypothesis.

Furthermore, the likelihood ratio is inherently linked to information
theoretic quantities and divergence measures. Assuming that \(Q\) is a
discrete distribution with support \(\chi\), one can write the negative
log-likelihood ratio as (compare Cover and Thomas, 2006)

\begin{align*}
-\log(LR_T(\theta_{H_0}|| \theta_{H_1}) & = & \sum_{i=1}^T \log(\frac{q(x_s|\theta_{H_0})}{ q(x_s|\theta_{H_1})}) \\
& = & \sum_{a \in \chi} T q(a|\hat{\theta}) \log(\frac{q(a|\theta_{H_0})}{ q(a|\theta_{H_1})}) \\
& = & \sum_{a \in \chi} T q(a|\hat{\theta}) \log(\frac{q(a|\theta_{H_0})}{ q(a|\theta_{H_1})}\frac{q(a|\hat{\theta})}{ q(a|\hat{\theta})}) \\
& = & T \KL(q(\hat{\theta})||q(\theta_{H_0})) - T \KL(q(\hat{\theta})||q(\theta_{H_1})),
\end{align*}

where \(\hat{\theta}\) is the maximum likelihood estimate of the
parameter \(\theta\) given the sample \(x\). If we hold the \(H_0\)
hypothesis fixed and set \(\theta_{H_1} = \hat{\theta}\) as often done
in praxis, then with \(\KL(Q(\hat{\theta})||Q(\hat{\theta})) = 0\):

\[
-\log(LR_T(\theta_{H_0}|| \theta_{H_1})) = -\log(LR_T(\theta_{H_0}|| \hat{\theta})) = \KL(Q(\hat{\theta}||Q(\theta_{H_0}))
\]

Furthermore, we have for any distributions \(Q_{H_0}\) and \(Q_{H_1}\)
for a sample \(x\) sampled from the \(H_0\) distribution, that

\[
\frac{1}{n} \log(LR_T(\theta_{H_0}||\theta_{H_1})) \rightarrow \KL(Q(\theta_{H_0})||Q(\theta_{H_1})) \quad \text{in probability},
\]

by the law of large numbers (or asymptotic equipartition property) for
the Kullback-Leibler divergence as in Theorem 11.8.1 in Cover and Thomas
(2006). Lastly, based on the characteristics shown before, for
frequentist tests the Kullback-Leibler divergence between
\(Q(\theta_{H_0})\) and \(Q(\theta_{H_1})\) gives the best possible
exponent for the error of the second type given a fixed error of the
first type. This is the result of the Chernoff-Stein Lemma (Theorem
11.8.3, Cover and Thomas, 2006), which is reminiscient of the complexity
\(\kappa_B\) derived by Kaufmann et al. (2016) and as described in
\autoref{chap:KullbackLeiblerPureExploration}:

\begin{theorem}[Chernoff-Stein Lemma] \label{theorem:ChernoffStein}
Let $X_1, X_2, ..., X_n$ be i.i.d. distributed according to $Q$. Consider the hypothesis test between two alternatives, $Q=P_1$ and $Q=P_2$, where $\KL(P_1||P_2) < \infty$. Let $A_n \subseteq \chi^n$ be an acceptance region for hypothesis $H_1$. Let the probabilities of error be $\alpha_n = P_1^n(A_n^C)$, and $\beta_n = P_2^n(A_n)$, and for $0< \epsilon <1/2$, define $\beta_n^{\epsilon} = \min_{A_n \subseteq \chi^n, \alpha_n < \epsilon} \beta_n$. Then

$$
\text{lim}_{n\rightarrow \infty} \frac{1}{n} \log \beta_n^{\epsilon} = - \KL(P_1||P_2)
$$
\end{theorem}

\subsubsection{Likelihood Ratio as Index in Sequential
Sampling}\label{likelihood-ratio-as-index-in-sequential-sampling}

We now employ the concept of the likelihood-ratio test for a new
sequential sampling strategy for the thresholding bandit problem. In
contrast to the cumulative regret problem or the best arm identification
problem, the thresholding bandit problem offers a natural \(H_0\)
hypothesis for the arms' distributions: the threshold \(\tau\). Given
that we want to classify the arms against the threshold, and given that
\(\tau\) is known and fixed, we can set a null hypothesis that is a
function of \(\tau\). In the simplest case, the distributions are
parameterized by their mean \(\theta_i = \mu_i\), and we set
\(\theta_{i,H_0} = \tau\) for all arms. The likelihood ratio then
quantifies how ``significant'' of a difference there is between the
threshold \(\tau\) and the current sample mean \(\hat{\mu}_i(t)\) of an
arm, assuming distributions for which the maximum likelihood estimate of
the mean is given by the empirical mean. In these cases, the likelihood
ratio will be smaller than 1 and decrease in sample size even if the
threshold and sample mean are constant.

Due to these characteristics, the likelihood ratio can serve as a
natural index for sampling strategies. Consider the setting described in
\autoref{sec:Setup} and \autoref{sec:TheAPTAlgorithm} though without
assuming \(R\)-sub-Gaussian distributions. Consider \(B_i^{APT}\) and
compare it to

\[
B_i^{LR}(t+1) = LR(\tau|| \hat{\mu}_i(t)) = \prod_{i=1}^T\frac{q(x_i|\tau)}{ q(x_i|\hat{\mu}_i(t))}.
\]

We have \(B_i(t) \leq 1\), \(t \uparrow \rightarrow B_i(t) \downarrow\),
and \(|\hat{\mu}_i(t) - \tau| \uparrow \rightarrow B_i(t) \downarrow\).
Consequently, by choosing the arm that maximizes \(B_i(t)\), we choose
an arm for which we either have not collected many samples, or for which
the sample mean is close to the threshold \(\tau\). In this sense, a
strategy based on the likelihood ratio is reminiscient of the index
employed by the APT algorithm. Indeed, assuming the samples are normally
distributed with the same variance for the distribution \(\nu_i\) under
both \(H_0\) and \(H_1\), then \(-\log(LR(\tau||\hat{\mu}_i(t)))\)
equals the index in APT. Why then would one use the likelihood based
strategy?

Given that the APT index and the likelihood-ratio based index are only
equivalent for normally distributed random variables with fixed standard
deviation, the new strategy promises efficiency improvements whenever
the true distribution is known and different from the normal
distribution. The most important example are Bernoulli arms with small
means as motivated. While the likelihood ratio can in theory be used for
any distributions, and arms could even have different classes of
distributions, the class of distribution has to be known upfront for
every arm. Given the role of the threshold \(\tau\) and the goal of
classification of the mean against \(\tau\), the likelihood ratio
strategy is arguably most naturally employed on distributions
parameterized by their mean. In the following descriptions, we thus
focus on univariate distributions of the exponential family.
Additionally, we also propose a version for the Normal distribution with
unknown mean. Before we describe the algorithm, we show that the
likelihood ratio has a very convenient form for this family of
distributions.

\subsubsection{Likelihood Ratio for Univariate Exponential Family of
Distributions}\label{likelihood-ratio-for-univariate-exponential-family-of-distributions}

We now focus on a thresholding bandit problem where the arms are
distributed according to \(\nu_{\theta_1}, ..., \nu_{\theta_K}\), which
are all from the same canonical exponential family of distributions with
scalar parameter \(\theta_i \in \Theta\). The probability density
function of the absolutely continuous distribution \(\nu_{\theta_i}\)
with regard to a dominating measure \(\rho\) on \(\mathbb{R}\) is given
by \(q(x|\theta) = h(x) \exp(T(x)\theta - b(\theta))\). As described in
Cappé et al. (2013), citing Lehmann and Casella (1998), the derivative
\(b'(\theta)\) of \(b(\theta)\) is an increasing continuous function
such that the expected value \(\mathbb{E}(\nu_{\theta}) = b'(\theta)\)
for all \(\theta \in \Theta\). Given that \(b'(x)\) is strictly convex,
we have a one-to-one relation between \(b'(\theta)\) and
\(\mathbb{E}(\nu_{\theta})\) through the continuous inverse
\(b'^{-1}(\theta)\). Consequently, one can just as well parameterize the
distribution its mean \(\mu = \mathbb{\nu_{\theta}}\) and write
\(\nu_{b'^{-1}(\mu)}\) with the probability density function (compare
Cappé et al., 2013)

@@

As we will see now, this statistic takes a very convenient closed form
for many common distributions.

\textbf{Example 1: Bernoulli Distribution}

@@

\begin{align*}
-\log(LR(\mu_0||\mu_1)) & = & n(\log\Big(\frac{1-\mu_0}{1-\mu_1}\Big)) - \Big(\log(\frac{\mu_0}{1-\mu_0}) - \log(\frac{\mu_1}{1-\mu_1})\Big)\sum_{i=1}^{n}x_i \\
& = & n \Big( (1-\hat{\mu})\log\Big(\frac{1-\mu_1}{1-\mu_0}\Big) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}) \Big)
\end{align*}

We observe that, if we let \(\mu_0 = \tau\) and \(\mu_1 = \hat{\mu}\)
(compare \autoref{sec:LRTests}), then

@@

\textbf{Example 2: Poisson Distribution}

@@

We observe that, if we let \(\mu_0 = \tau\) and \(\mu_1 = \hat{\mu}\)
(compare \autoref{sec:LRTests}), then

@@

\textbf{Example 3: Normal distributions with known variance
\(\sigma^2\)}

@@

And again, even though the distribution is continuous (compare
\autoref{sec:LRTests}), if we let \(\mu_1 = \hat{\mu}\) and
\(\mu_0 = \tau\), then

@@

\textbf{Example 4: Gamma distributions with known shape parameter
\(\alpha\)}

@@

\[
-\log(LR(\mu_0||\mu_1)) = n\alpha(-\log(\frac{\alpha}{\mu_0}) - \log(\frac{\mu_1}{\alpha})) - (-\frac{\alpha}{\mu_0} + \frac{\alpha}{\mu_1})\sum_{i=1}^{n}x_i = n\alpha \Big(\log\big(\frac{\mu_0}{\mu_1}\big) - \frac{\hat{\mu}}{\mu_1} + \frac{\hat{\mu}}{\mu_0}\Big).
\]

And again, even though the distribution is continuous (compare
\autoref{sec:LRTests}), if we let \(\mu_1 = \hat{\mu}\) and
\(\mu_0 = \tau\), then

@@

Given these results for the exponential family of distributions, we can
now define the Simple Likelihood Ratio algorithm.

\subsection{\texorpdfstring{1D-Simple Likelihood Ratio Algorithm
\label{sec:1DSLR}}{1D-Simple Likelihood Ratio Algorithm }}\label{d-simple-likelihood-ratio-algorithm}

Consider again the setting in \autoref{sec:Setup} though without the
assumption of \(R\)-sub-Gaussian distributions. Instead, assume that the
class of distributions that the arms \(i \in \mathbb{A}\) follow is
known upfront. For the 1D-SLR algorithm, we restrict the potential
classes to the univariate exponential family of distributions and write
the class generally as \(\nu(\mu, \theta)\). Here, \(\theta\) describes
additional parameters known upfront, as for example
\(\sigma_1, ..., \sigma_K\) for Normal distributions, or
\(\alpha_1, ..., \alpha_K\) for Gamma distributions. Additionally to
\(\nu(\mu)\), the 1D-SLR algorithm receives as input the definition of
the problem \((\tau, \epsilon)\) (compare Locatelli et al., 2016). As
initialization, the algorithm pulls each arm once. At time \(t>K\), the
algorithm updates the number of pulls for arm \(i\) at time \(t\),
\(T_i(t)\), for each arm, as well as the empirical mean
\(\hat{\mu}_i(t)\) of arm \(i\) after \(T_i(t)\) pulls. The empirical
mean is calculated for each arm \(k\in \{1,...,K\}\) as
\(\hat{\mu}_i(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)} x_{i,s}\), with
\(T_i(t) = \sum_{s=1}^{t}\mathbb{I}(I_s = i)\). It serves as maximum
likelihood estimate of the arm's mean. As in Locatelli et al. (2016),
the \(X_{i,s}\) is the sample received when pulling arm \(i\) for the
\(s\)-th time. Depending on the distribution, the algorithm then
computes the index
